---
title: "Interactive Data Visualization with Plotly"
author: "Phuoc Nguyen"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  ioslides_presentation: default
  beamer_presentation:
    theme: default
    colortheme: default
    fonttheme: default
---

## Welcome to Interactive Data Visualization

This presentation demonstrates the power of Plotly for creating interactive visualizations.

---

## What is Plotly?

- **Interactive plotting library** for R
- Creates **web-based visualizations**
- Supports **hover effects** and **zoom**
- Perfect for **data exploration**

---

## Sample Dataset

Let's explore the **mtcars** dataset:

```{r setup, include=FALSE}
library(plotly)
library(dplyr)
```

```{r echo=FALSE}
# Display first few rows
head(mtcars)
```

---

## Interactive Scatter Plot

```{r}
# Create interactive scatter plot
p <- plot_ly(data = mtcars, 
             x = ~wt, 
             y = ~mpg, 
             color = ~cyl,
             size = ~hp,
             text = ~paste("Car: ", rownames(mtcars), 
                          "<br>Weight: ", wt, 
                          "<br>MPG: ", mpg,
                          "<br>Cylinders: ", cyl,
                          "<br>Horsepower: ", hp),
             hoverinfo = "text") %>%
  add_markers() %>%
  layout(title = "Car Performance Analysis",
         xaxis = list(title = "Weight (1000 lbs)"),
         yaxis = list(title = "Miles per Gallon"))

p
```

---

## Key Insights

- **Lighter cars** tend to have **better fuel efficiency**
- **More cylinders** generally mean **higher weight**
- **Horsepower** varies significantly across car types

---

## Interactive Bar Chart

```{r}
# Create interactive bar chart
avg_mpg <- mtcars %>%
  group_by(cyl) %>%
  summarise(avg_mpg = mean(mpg), .groups = 'drop')

p2 <- plot_ly(data = avg_mpg,
              x = ~cyl,
              y = ~avg_mpg,
              type = "bar",
              text = ~paste("Cylinders: ", cyl, "<br>Avg MPG: ", round(avg_mpg, 2)),
              hoverinfo = "text") %>%
  layout(title = "Average MPG by Number of Cylinders",
         xaxis = list(title = "Number of Cylinders"),
         yaxis = list(title = "Average Miles per Gallon"))

p2
```

---

## 3D Visualization

```{r}
# Create 3D scatter plot
p3 <- plot_ly(data = mtcars,
               x = ~wt,
               y = ~mpg,
               z = ~hp,
               color = ~cyl,
               text = ~paste("Car: ", rownames(mtcars)),
               hoverinfo = "text") %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = "Weight"),
                      yaxis = list(title = "MPG"),
                      zaxis = list(title = "Horsepower")))

p3
```

---

## Benefits of Interactive Visualization

- **Zoom and pan** for detailed exploration
- **Hover effects** for instant data access
- **Multiple views** of the same data
- **Professional presentation** quality

---

## Thank You!

**Interactive visualizations** make data more engaging and accessible.

**Plotly** provides the tools to create stunning, interactive charts.

---

## Contact Information

- **Author:** Phuoc Nguyen
- **Date:** `r format(Sys.Date(), '%Y-%m-%d')`
- **Technology:** R Markdown + Plotly
