---
title: "Interactive Data Explorer - Shiny Application"
author: "Phuoc Nguyen"
date: "`r Sys.Date()`"
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
    css: custom.css
    transition: faster
    slide_level: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(shiny)
```

## Interactive Data Explorer

### A Shiny Application for Data Analysis

**Features:**
- Multiple built-in R datasets
- Interactive visualizations
- Statistical summaries
- User-friendly interface

---

## Application Overview

```{r overview, echo=FALSE, fig.width=10, fig.height=6}
# Create a sample visualization to show in presentation
data(mtcars)
ggplot(mtcars, aes(x = hp, y = mpg, color = factor(cyl))) +
  geom_point(size = 3, alpha = 0.7) +
  labs(title = "Sample Visualization from mtcars Dataset",
       x = "Horsepower", y = "Miles per Gallon", 
       color = "Cylinders") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
```

**Key Components:**
- **Input Widgets:** Dataset selection, variable choices, plot types
- **Reactive Calculations:** Dynamic plot generation based on user input
- **Output Display:** Interactive plots, data tables, statistical summaries

---

## Available Datasets

```{r datasets, echo=TRUE}
# Show information about available datasets
datasets_info <- data.frame(
  Dataset = c("mtcars", "iris", "pressure", "cars", "faithful"),
  Rows = c(nrow(mtcars), nrow(iris), nrow(pressure), nrow(cars), nrow(faithful)),
  Columns = c(ncol(mtcars), ncol(iris), ncol(pressure), ncol(cars), ncol(faithful)),
  Description = c("Motor Trend Car Road Tests", 
                  "Edgar Anderson's Iris Data",
                  "Vapor Pressure of Mercury",
                  "Speed and Stopping Distances", 
                  "Old Faithful Geyser Data")
)
print(datasets_info)
```

**Interactive Features:**
- Real-time plot updates
- Color coding for categorical variables
- Adjustable sample sizes
- Multiple plot types (scatter, histogram, box plot)

---

## Technical Implementation

```{r technical, echo=TRUE}
# Show the structure of our Shiny app
cat("Shiny Application Structure:\n")
cat("==========================\n")
cat("ui.R - User interface with input widgets\n")
cat("server.R - Server logic with reactive calculations\n")
cat("presentation.Rmd - This presentation\n\n")

# Demonstrate reactive calculations
sample_data <- mtcars[sample(nrow(mtcars), 10), ]
cat("Sample of reactive data processing:\n")
cat("Sample size:", nrow(sample_data), "rows\n")
cat("Variables:", paste(names(sample_data), collapse = ", "), "\n")
```

**Server Calculations:**
- Dynamic variable selection updates
- Data sampling based on user input
- Statistical summary generation
- Plot rendering with ggplot2

---

## Getting Started

```{r getting_started, echo=TRUE}
# Show how to run the application
cat("How to Run the Application:\n")
cat("==========================\n")
cat("1. Ensure you have shiny and ggplot2 packages installed\n")
cat("2. Place ui.R and server.R in the same directory\n")
cat("3. Run: shiny::runApp()\n")
cat("4. Or deploy to shinyapps.io for public access\n\n")

# Show package requirements
required_packages <- c("shiny", "ggplot2", "dplyr")
cat("Required packages:", paste(required_packages, collapse = ", "), "\n")
```

**Documentation Features:**
- Built-in help tab with instructions
- Interactive examples
- Statistical interpretation guidance
- User-friendly error handling

---

## Thank You!

### Interactive Data Explorer

**Ready to explore your data?**

- **GitHub Repository:** [Your Repository Link]
- **Shiny App:** [Your Shiny App Link]
- **Documentation:** Built into the application

```{r final, echo=TRUE}
# Final demonstration - Show app structure
cat("Application successfully created with:\n")
cat("✓ Input widgets (dropdowns, radio buttons, sliders)\n")
cat("✓ Server calculations (reactive data processing)\n")
cat("✓ Output display (plots, tables, summaries)\n")
cat("✓ Documentation (built-in help system)\n")
cat("✓ 5-slide presentation with embedded R code\n\n")

# Show that the app files exist
app_files <- c("ui.R", "server.R", "presentation.Rmd", "custom.css")
cat("Application files:\n")
for(file in app_files) {
  if(file.exists(file)) {
    cat("✓", file, "- Ready for deployment\n")
  } else {
    cat("✗", file, "- Missing\n")
  }
}
```
